---
title: "P8105 Homework 1"
author: "Riyadh Baksh"
date: "September 21, 2024"
output: github_document
---

# Problem 1

```{r, message=FALSE}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

The following are the variables within the `penguins` dataset.

* `species`: either "Adelie", "Chinstrap", or "Gentoo"
* `island`: either "Biscoe", "Dream", or "Torgersen"
* `bill_length_mm`: bill length in mm
* `bill_depth_mm`: bill depth in mm
* `flipper_length_mm`: flipper length in mm
* `body_mass_g`: body mass in grams
* `sex`: either male or female
* `year`: either 2007, 2008, or 2009

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

```{r}
flipper_length = penguins$flipper_length_mm
mean_flipper_length = mean(flipper_length, na.rm = TRUE)
```

The mean flipper length is `r round(mean_flipper_length, 2)` mm.

The scatterplot of `flipper_length_mm` versus `bill_length_mm` is shown below, with points colored by `species`.

```{r, message=FALSE, warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("penguins.png")
```

# Problem 2

The dataframe is created below.

```{r, message=FALSE}
library(tidyverse)
variable_df = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = c("a","b","c","d","e","f","g","h","i","j"),
  vec_factor = factor(c("low", "medium", "low", "medium", "high", "high", "low", "low", "medium", "medium"))
)

show(variable_df)
```
The means of each variable is taken below. The mean is correctly calculated for `vec_numeric` since these are numbers, and is correctly calculated for `vec_logical` since `TRUE` and `FALSE` are coded as 1 and 0, respectively. However, the mean is not calculated for `vec_char` and `vec_factor` since these are stored as text and the `mean()` function requires numeric values.

```{r}
mean(pull(variable_df, vec_numeric))
```


```{r}
mean(pull(variable_df, vec_logical))
```


```{r}
mean(pull(variable_df, vec_char))
```

```{r}
mean(pull(variable_df, vec_factor))
```

If the `as.numeric()` function is applied, then `vec_logical` is correctly understood as either 1 or 0. For `vec_char`, the function results in `NA` since it is not defined what a character represents numerically. For `vec_factor`, the function results in unique numbers corresponding to each factor, although there is not a apparent order or pattern to how the numbers are corresponded, again since R does not understand the ordered meaning of text. These results explain why the mean function worked well for the logical vector, but couldn't work for the character or factor vectors.

```{r, eval=FALSE}
as.numeric(pull(variable_df, vec_logical))
```

```{r, eval=FALSE, message=FALSE}
as.numeric(pull(variable_df, vec_char))
```

```{r, eval=FALSE}
as.numeric(pull(variable_df, vec_factor))
```
