---
title: "P8105 Homework 1"
author: "Riyadh Baksh"
date: "September 21, 2024"
output: github_document
---

# Problem 1

```{r, message=FALSE}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

The following are the variables within the `penguins` dataset.

* `species`: either "Adelie", "Chinstrap", or "Gentoo"
* `island`: either "Biscoe", "Dream", or "Torgersen"
* `bill_length_mm`: bill length in mm
* `bill_depth_mm`: bill depth in mm
* `flipper_length_mm`: flipper length in mm
* `body_mass_g`: body mass in grams
* `sex`: either male or female
* `year`: either 2007, 2008, or 2009

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

```{r}
flipper_length = penguins$flipper_length_mm
mean_flipper_length = mean(flipper_length, na.rm = TRUE)
```

The mean flipper length is `r round(mean_flipper_length, 2)` mm.

The scatterplot of `flipper_length_mm` versus `bill_length_mm` is shown below, with points colored by `species`.


```{r, message=FALSE, warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("penguins.png")
```


# Problem 2


